### 2015-04-16 nir_maps.tex
maps_latex

\documentclass{article}
\usepackage{graphicx}

\begin{document}

\title{Maps of the proportion vaccinated in area units using NIR data}
\author{David Hayman}

\maketitle
\newpage
\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2006.pdf}
    \caption{MMR1 2006}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2006.pdf}
    \caption{MMR2 2006}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2007.pdf}
    \caption{MMR1 2007}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2007.pdf}
    \caption{MMR2 2007}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2008.pdf}
    \caption{MMR1 2008}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2008.pdf}
    \caption{MMR2 2008}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2009.pdf}
    \caption{MMR1 2009}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2009.pdf}
    \caption{MMR2 2009}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2010.pdf}
    \caption{MMR1 2010}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2010.pdf}
    \caption{MMR2 2010}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2011.pdf}
    \caption{MMR1 2011}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2011.pdf}
    \caption{MMR2 2011}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2012.pdf}
    \caption{MMR1 2012}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2012.pdf}
    \caption{MMR2 2012}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2013.pdf}
    \caption{MMR1 2013}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2013.pdf}
    \caption{MMR2 2013}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR1_NIR_2014.pdf}
    \caption{MMR1 2014}
\end{sidewaysfigure}

\begin{sidewaysfigure}
    \centering
    \includegraphics[width=5.0in]{nir_census_MMR2_NIR_2014.pdf}
    \caption{MMR2 2014}
\end{sidewaysfigure}

\end{document}